Reproducible Research: Peer Assessment 1(updated Version)
========================================

## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day  
- Use aggregate to get total number of steps per day and store in steps.date.  
- Make a barplot using the "steps" part of steps.date, use the date part for bar labels.  
```{r}
steps.date <- aggregate(steps ~ date, data = activity, FUN = sum)                                                                                                                                                                                                         
barplot(steps.date$steps, names.arg = steps.date$date, xlab = "date", ylab = "steps")
```


2. Calculate and report the **mean** and **median** total number of
   steps taken per day
- Take mean of the steps column of steps.date
- Take median of the steps column of steps.date
```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute
   interval (x-axis) and the average number of steps taken, averaged
   across all days (y-axis)
- Use aggregate to take the mean of all the steps for a certain interval, store in steps.interval
- Plot on line chart
```{r}
steps.interval <- aggregate(steps ~ interval, data=activity, FUN=mean)
plot(steps.interval, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset,contains the maximum number of steps?
- Retreieve the row in the interval column where the maximum number of steps took place using which.max
```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the
   dataset (i.e. the total number of rows with `NA`s)
- Take the sum of all the NAs in the dataset with is.na
```{r}
sum(is.na(activity))
```

2. Devise a strategy for filling in all of the missing values in the
   dataset. The strategy does not need to be sophisticated. For
   example, you could use the mean/median for that day, or the mean
   for that 5-minute interval, etc.

*We will use the means for the 5-minute intervals as fillers for missing values.*

3. Create a new dataset that is equal to the original dataset but with
   the missing data filled in.
- First use merge to join the activity and steps.interval with the interval column, and to modify the new steps column with a .y suffix
- Then find out all rows with NA values in steps column
- Replace those NA values with the appropriate mean values
- Remove the steps.y column
```{r}
activity <- merge(activity, steps.interval, by = "interval", suffixes = c("", ".y"))
nas <- is.na(activity$steps)
activity$steps[nas] <- activity$steps.y[nas]
activity <- activity[, c(1:3)]
```

4. Make a histogram of the total number of steps taken each day and
   Calculate and report the **mean** and **median** total number of
   steps taken per day. Do these values differ from the estimates from
   the first part of the assignment? What is the impact of imputing
   missing data on the estimates of the total daily number of steps?
- Use aggregate to get total number of steps per day and store in steps.date.  
- Make a barplot using the "steps" part of steps.date, use the date part for bar labels.  
- Take mean of the steps column of steps.date
- Take median of the steps column of steps.date
```{r}   
steps.date <- aggregate(steps ~ date, data = activity, FUN = sum)
barplot(steps.date$steps, names.arg = steps.date$date, xlab = "date", ylab = "steps")

mean(steps.date$steps)

median(steps.date$steps)
```

*The impact of the missing data seems rather low, at least when estimating the total number of steps per day.*


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels --
   "weekday" and "weekend" indicating whether a given date is a
   weekday or weekend day.
- Create a function that takes a date, and through the weekdays function, will set the date's factor as weekend or weekday
- Apply the function to the activity data frame
```{r, cache=TRUE}
daytype <- function(date) {
        if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
                "weekend"
        } else {
                "weekday"
        }
}
activity$daytype <- as.factor(sapply(activity$date, daytype))
```

2. Make a panel plot containing a time series plot (i.e. `type = "l"`)
   of the 5-minute interval (x-axis) and the average number of steps
   taken, averaged across all weekday days or weekend days
   (y-axis).
- Set the panels to 2 by 1 (one for weekdays and one for weekends)
- For all elements with a weekend factor,
    - Split up all of the steps by the 5 minute intervals, and take the mean of these splits
    - Plot all of the intervals by their average number of steps
- Do the same for elements with a weekday factor
```{r}
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
        steps.type <- aggregate(steps ~ interval, data = activity, subset = activity$daytype == type, FUN = mean)
        plot(steps.type, type = "l", main = type)
}
```